---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data

```{r, echo=TRUE}
library(ggplot2)
library(dplyr)
if(!file.exists("./data")) {dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = "./data/activity.zip")
projectData <- unzip("./data/activity.zip", exdir = "./data" )
View(projectData)
df <- read.csv("./data/activity.csv")
View(df)
str(df)
df$date <- as.Date(df$date, format = "%Y-%m-%d")
summary(df)
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r, echo=TRUE}
stepsbyday <- aggregate(steps ~ date, df, sum, na.rm = TRUE)
head(stepsbyday)
```

2. Make a histogram of the total number of steps taken each day

```{r hist, echo=TRUE}
 hist(stepsbyday$steps)
```

3. Calculate and report the mean and median of the total number of steps taken per day

    Here's the mean of the total number of steps
    
```{r, echo=TRUE}
meanstepsbyday <- mean(stepsbyday$steps)
meanstepsbyday
```
    Here's the median of the total number of steps 
    
```{r, echo=TRUE}
medianstepsbyday <- median(stepsbyday$steps)
medianstepsbyday
```

## What is the average daily activity pattern?
1. Make a time series plot

```{r plot, echo=TRUE}
stepsperinterval <- aggregate(steps ~ interval, data = df, mean, na.rm = TRUE)
with(stepsperinterval, plot(steps~interval, type = "l", xlab = " 5-minute interval", ylab = "average number of steps taken"))
```

2. Maximum number of steps

```{r, echo=TRUE}
maxsteps <- stepsperinterval[which.max(stepsperinterval$steps),]$interval
maxsteps
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset

```{r, echo=TRUE}
totalmissingvalues <- sum(is.na(df$steps))
totalmissingvalues
```

2. Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r, echo=TRUE}
        newdf <- df
        na <- is.na(df$steps)
        avg_interval <- with(newdf, tapply(steps, interval, mean, na.rm = TRUE, simplify = TRUE))
        newdf$steps[na] <- avg_interval[as.character(newdf$interval[na])]

```

3. Calculating the steps

```{r, echo=TRUE}
newsteps <- newdf %>%
    filter(!is.na(newdf$steps)) %>%
    group_by(date) %>%
    summarise(steps = sum(steps)) %>%
    print
    ```
4. Make a histogram of the total number of steps taken each day

```{r, echo=TRUE}
        hist(newsteps$steps, xlab = "steps per day", main = "total number of steps each day")
```

5.Calculate and report the mean and median total number of steps taken per day
```{r, echo=TRUE}
meanNewsteps <- mean(newsteps$steps, na.rm = TRUE)
meanNewsteps
medianNewsteps <- median(newsteps$steps, na.rm = TRUE)
medianNewsteps
```
## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
newdf$weekday <- as.factor(weekdays(newdf$date))
newdf$weektype <- if_else(newdf$weekday %in% c("Saturday", "Sunday"), "weekend", "weekday")
head(newdf)
```
2. Make a panel plot containing a time series plot
```{r, echo=TRUE}
 newInterval <- aggregate(steps ~ interval + weektype, data = newdf, mean, na.rm = TRUE)
 g <- ggplot(newInterval, aes(x = interval, y = steps, color = weektype))
 g <- g + geom_line() + facet_wrap(~ weektype, ncol = 1, nrow = 2)
 print(g)
 ```
 
